# Data transformation

For cleaning the data, we mainly used two functions: clean_tweets and tagging. 

clean_tweets function removes the parts of the tweets that are not relevant to sentiment analysis. For instance, we removed URLs, @username, special characters, newlines, and whitespaces to have the texts ready for analysis.

tagging function selects the tweets that have at least one of the tags related to COVID-19. For instance, a tweet with "corona" or "covid-19" would be kept for analysis, while a tweet that has none of the tags in the list would be excluded.

We used these two functions on all our datasets for four groups to work with only cleaned tweets that are related to COVID-19.

```{r, include=FALSE}
knitr::opts_chunk$set(echo= TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(patchwork)
library(ggnewscale)
library(rtweet)
library(sjmisc)
library(readr)
library(rtweet)
```

## Functions for data cleaning
```{r}
clean_tweets <- function(x) {
  x %>%
    # Remove URLs
    str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
    # Remove mentions e.g. "@my_account"
    str_remove_all("@[[:alnum:]_]{4,}") %>%
    # Replace "&" character reference with "and"
    str_replace_all("&amp;", "and") %>%
    # Remove punctuation, using a standard character class
    str_remove_all("[[:punct:]]") %>%
    # Remove "RT: " from beginning of retweets
    str_remove_all("^RT:? ") %>%
    # Replace any newline characters with a space
    str_replace_all("\\\n", " ") %>%
    # Make everything lowercase
    str_to_lower() %>%
    # Remove any trailing whitespace around the text
    str_trim("both") %>% 
    # remove unnecessary space
    str_replace_all("  "," ")
}
```

```{r}
tag <- c('corona', '#corona', 'coronavirus', '#coronavirus', 'covid', '#covid', 'covid19', '#covid19', 'covid-19', '#covid-19', 'sarscov2', '#sarscov2', 'sars cov2', 'sars cov 2', 'covid_19', '#covid_19', '#ncov', 'ncov', '#ncov2019', 'ncov2019', '2019-ncov', '#2019-ncov', 'pandemic', '#pandemic #2019ncov', '2019ncov', 'quarantine', '#quarantine', 'flatten the curve', 'flattening the curve', '#flatteningthecurve', '#flattenthecurve', 'hand sanitizer', '#handsanitizer', '#lockdown', 'lockdown', 'social distancing', '#socialdistancing', 'work from home', '#workfromhome', 'working from home', '#workingfromhome', 'ppe', 'n95', '#ppe', '#n95', '#covidiots', 'covidiots', 'herd immunity', '#herdimmunity', 'pneumonia', '#pneumonia', 'chinese virus', '#chinesevirus', 'wuhan virus', '#wuhanvirus', 'kung flu', '#kungflu', 'wearamask', '#wearamask', 'wear a mask', 'vaccine', 'vaccines', '#vaccine', '#vaccines', 'corona vaccine', 'corona vaccines', '#coronavaccine', '#coronavaccines', 'face shield', '#faceshield', 'face shields', '#faceshields', 'health worker', '#healthworker', 'health workers', '#healthworkers', '#stayhomestaysafe', '#coronaupdate', '#frontlineheroes', '#coronawarriors', '#homeschool', '#homeschooling', '#hometasking', '#masks4all', '#wfh', 'wash ur hands', 'wash your hands', '#washurhands', '#washyourhands', '#stayathome', '#stayhome', '#selfisolating', 'self isolating')

tagging <- function(df, cleaned_text) {
  output <- c()
  for (words in str_split(cleaned_text, " ")) {
    f <- if_else(any(words %in% tag), 1, 0)
    output <- append(output, f)
  }
  df$ouput <- output
  filtered_df <- df %>% filter(output ==1) 
  return(filtered_df)
}
```

## Politicians
```{r}
df3 <- read_csv("./resources/data_politicians.csv")
cleaned_text <- clean_tweets(df3$text)
data_politicians_filtered <- tagging(df3, cleaned_text)
write_csv(data_politicians_filtered, "./resources/data_politicians_filtered.csv")
```

## Medical Associations
```{r}
df2 <- read_csv("./resources/data_newsmedia.csv")
cleaned_text2 <- clean_tweets(df2$text)
data_newsmedia_filtered <- tagging(df2, cleaned_text2)
write_csv(data_newsmedia_filtered, "./resources/data_newsmedia_filtered.csv")
```

## Celebrities
```{r}
df4 <- read_csv("./resources/data_celebrities.csv")
cleaned_text4 <- clean_tweets(df4$text)
data_celebrities_filtered <- tagging(df4, cleaned_text4)
write_csv(data_celebrities_filtered, "./resources/data_celebrities_filtered.csv")
```
