# Interactive component

```{r}
library(base)
```

```{r}
df <- read_csv("./resources/data_combined_filtered_sentiment.csv")
```

```{r}
df_t <- df %>% select(created_at, group, sentiment_score)
```

```{r}
times <- as.character(df_t$created_at)

dates <- c()

for (time in times){
  date <- strsplit(time, " +")[[1]][1]
  dates <- c(dates, date)
}

df_t$created_at <- dates
df_t
```
We made two csv files for interactive graph. Both csv contain four columns - date, politicians' sentiment scores, medicals' sentiment scores, celebrities' sentiment scores. interactive_with_na.csv contains all dates that at least one of the groups has sentiment scores, which means there are NAs in some entries for each group. interactive_without_na.csv only contains dates that all three groups have sentiment scores.

```{r}
df_p <- df_t %>% filter(group == "politicians") 
df_m <- df_t %>% filter(group == "medical")
df_c <- df_t %>% filter(group == "celebrities")

df_p <- df_p %>%
  group_by(created_at) %>%
  summarise(sentiment_score = mean(sentiment_score))

df_m <- df_m %>%
  group_by(created_at) %>%
  summarise(sentiment_score = mean(sentiment_score))

df_c <- df_c %>%
  group_by(created_at) %>%
  summarise(sentiment_score = mean(sentiment_score))

df_combined <- merge(df_p, df_m, by = "created_at")
df_combined <- merge(df_combined, df_c, by = "created_at")
colnames(df_combined) <- c('x', 'politicians', 'medicals', 'celebrities')

df_combined_na <- merge(df_p, df_m, by = "created_at", all = TRUE)
df_combined_na <- merge(df_combined_na, df_c, by = "created_at", all = TRUE)
colnames(df_combined_na) <- c('x', 'politicians', 'medicals', 'celebrities')

write_csv(df_combined_na, "./resources/interactive_with_na.csv")
write_csv(df_combined, "./resources/interactive_without_na.csv")
```
Using these csv files, we made an interactive graph. Using the drop-down button at top left, we can choose which group's sentiment score graph over time to display. We smoothed the curve (because there are many missing dates in the dataframe) by using d3 built-in function, d3.curveBasis. The graph will start by showing politicians' sentiment score graph by default, and choosing another group from the button will transform the graph accordingly. The benefit of using this interactive graph is that we can see each group's sentiment score changes more closely, and also, we can tell the difference between two groups by seeing how the graph transforms as we choose another group from the button. 

And as we can see from the graph, sentiment score of celebrities move in the similar direction as the medicals' score. One can go back and forth between the two groups' graphs to see that the general flow of the graphs don't change much, although the celebrities' scores are fluctuating more. But the politicians' score move in the opposite direction of the medical's score. One can go back and forth between the two groups' graphs to see that the general flow of the graphs change dramatically. 


<<insertHTML:[test.html]
```{r, echo=FALSE}
htmltools::includeHTML("interactive.html")
```









